---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---

Here, we want to identify the potential changes in phosphorylation which may occur following inhibition of EGFR as per the Orre et al (2018) 'SubCellBarCode' paper (https://www.cell.com/molecular-cell/pdfExtended/S1097-2765(18)31005-0), in which they used gefitinib to inhibit EGFR in a 746-750 del EGFR mutant cell line (HCC827).

In Zhang et al (2017; https://www.mcponline.org/content/16/5/891/tab-figures-data), they used EGFR inhibitors on a variety of backgrounds, including PC9 which is also 746-750 del. Unfotunately, they didn't use Gefitinib but rather erlotinib, which appears to be a similar first generation EGFR tyrosine kinase inhibitor (TKI) so this seems close enough.
```{r}
library(tidyverse)
```

```{r}
.data <- readxl::read_excel('../raw/Zhang_et_al_sup_data_1.xlsx', sheet=6, skip=1)
```
For PC9, the SILAC experiment was conducted such that L=Serum Starved, M=EGF, H=EGF+erlotinib. We are interested in the sites which are erlotinib-sensitive, e.g H/M.

In the paper, they state there are 73 erlotinib-sensitive phosohorylation sites. Le'ts inspect the H/M ratios to see if we can find these as a check we understand their data correctly.
```{r}
.data %>% ggplot(aes(log2(`Ratio H/M normalized`))) + geom_density() + theme_bw() +
  geom_vline(xintercept=-1, linetype=2)
```

```{r}
erlotinib_sensitive <- .data %>%
  filter(log2(`Ratio H/M normalized`)<(-1)) %>%
  arrange(`Ratio H/M normalized`)

print(nrow(erlotinib_sensitive))
print(head(erlotinib_sensitive))
```
OK, we have our 73 sites.

From here we could either take the top most sensitive and compare to the least sensitive, e.g subset like so. Note that it's possible for a protein to have sites which are both highly sensitive and highly insensitive so need to start by grouping sites by proteins and identifying the most sensitive site
```{r}
.data_protein <- .data %>%
  group_by(Protein) %>%
  summarise(`Ratio H/M normalized`=min(`Ratio H/M normalized`))

most_sensitive <- .data_protein %>% arrange(`Ratio H/M normalized`) %>%
  pull(Protein) %>% head(50)

least_sensitive <- .data_protein %>% arrange(`Ratio H/M normalized`) %>%
  pull(Protein) %>% tail(50)

.data_protein %>% mutate(sensitive=Protein %in% most_sensitive,
                         insensitive=Protein %in% least_sensitive,
                         label=recode(interaction(sensitive, insensitive),
                                      'FALSE.FALSE'='Other',
                                      'TRUE.FALSE'='Sensitive',
                                      'FALSE.TRUE'='Insensitive')) %>%
  ggplot(aes(log2(`Ratio H/M normalized`), fill=label)) +
  geom_histogram() +
  theme_bw() +
  scale_fill_discrete(name='') +
  xlab('EGFR-TKI vs EGF (log2)')
```



